---
title: "Creating a Linear Regression Model to Help Predict the Best Sales Price for Homes in Ames Iowa"
author: "Anika Abrahamson,Yaqiong(Juno) Cao, Dexter Nguyen, Michael Ruch, Xinying(Silvia) Sun"
date: "September 24, 2020"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message=FALSE, warning=FALSE)
```

<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
  }
td {  /* Table  */
  font-size: 8px;
}
h1.title {
  font-size: 38px;
  color: DarkRed;
}
h1 { /* Header 1 */
  font-size: 28px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: DarkBlue;
}

code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>

```{r}
library(dataQualityR)
library(e1071)
library(mice)
library(corrplot)
library(ggplot2)
library(ggpubr)
library(scales)
library(caret)
library(dplyr)
library(tidyverse)
library(sf)
library(gganimate)
library(MASS)
```

# Outline:
1. Introduction - Business Context
2. Data Preparation
2.1. Data collection
2.2. Data Cleaning
2.3. Data transformation
2.4. Correlation Analysis
3. Exploratory Data Analysis - EDA
4. Modeling
4.1. Modeling with top 10 variables
4.2. Modeling with all variables
5. Evaluation

# 1. Introduction

# 2. Data Preparation

## 2.1. Data collection

## 2.2. Data cleaning

- Data reading
```{r}
df<-read.delim("AmesHousing.txt")
dim(df)
```
- Data overview
```{r}
str(df)
summary(df)
```

- Remove 2 columns: Order and PID
```{r}
df$Order<-NULL # Order column looks like meaningless
df$PID<-NULL # PID column looks like meaningless
```

- change subclass from num to cat 
```{r}
class(df$MS.SubClass)
df$MS.SubClass
df$MS.SubClass = as.factor(df$MS.SubClass)
class(df$MS.SubClass)
levels(df$MS.SubClass)
```

- having a data report for cat and num 
```{r}
num.file <- paste(getwd(), "/dqames_num.csv", sep= "")
cat.file <- paste(getwd(), "/dqames_cat.csv", sep= "")
checkDataQuality(data= df, out.file.num= num.file, out.file.cat= cat.file)
```

- num - remove NAs
```{r}
df=df[which(is.na(df$Mas.Vnr.Area)==FALSE),]
df=df[which(is.na(df$BsmtFin.SF.1)==FALSE),]
df=df[which(is.na(df$BsmtFin.SF.2)==FALSE),]
df=df[which(is.na(df$Bsmt.Unf.SF)==FALSE),]
df=df[which(is.na(df$Total.Bsmt.SF)==FALSE),]
df=df[which(is.na(df$Garage.Yr.Blt)==FALSE),]
df=df[which(is.na(df$Garage.Cars)==FALSE),]
df=df[which(is.na(df$Garage.Area)==FALSE),]
df=df[which(is.na(df$Bsmt.Full.Bath)==FALSE),]
df=df[which(is.na(df$Bsmt.Half.Bath)==FALSE),]
```

- impute missing value by sampling 
```{r}
imputed = mice(df, seed = 169, method = "sample")
micedf = complete(imputed, 3)
df$Lot.Frontage = micedf$Lot.Frontage
```

- cat - remove Alley/pool/fence/misc.feature column
```{r}
df$Alley<-NULL
df$Pool.QC<-NULL
df$Fence<-NULL
df$Misc.Feature<-NULL
```

- cat - remove NAs for feature other cat culumns
```{r}
df=df[which(is.na(df$Bsmt.Qual)==FALSE),]
df=df[which(is.na(df$Bsmt.Cond)==FALSE),]
df=df[which(is.na(df$Bsmt.Exposure)==FALSE),]
df=df[which(is.na(df$BsmtFin.Type.1)==FALSE),]
df=df[which(is.na(df$BsmtFin.Type.2)==FALSE),]
df=df[which(is.na(df$Electrical)==FALSE),]
```

- replace the fireplace's NA with the most frequent level
```{r}
df$Fireplace.Qu[is.na(df$Fireplace.Qu)] <- "Gd" # I cannot found a better way to deal with the missing cat value 
```

- check current data report 
```{r}
num.file <- paste(getwd(), "/dqames_num.csv", sep= "")
cat.file <- paste(getwd(), "/dqames_cat.csv", sep= "")
checkDataQuality(data= df, out.file.num= num.file, out.file.cat= cat.file)
```

- Replace contents of two columns: Kitchen_Qual and Bsmt_Qual
```{r}
df$Kitchen.Qual <- as.character(df$Kitchen.Qual)
df$Kitchen.Qual[df$Kitchen.Qual == "Ex"] <- "Excellent"
df$Kitchen.Qual[df$Kitchen.Qual == "Gd"] <- "Good"
df$Kitchen.Qual[df$Kitchen.Qual == "TA"] <- "Typical"
df$Kitchen.Qual[df$Kitchen.Qual == "Fa"] <- "Fair"
df$Kitchen.Qual[df$Kitchen.Qual == "Po"] <- "Poor"
df$Kitchen.Qual <- as.factor(df$Kitchen.Qual)

df$Bsmt.Qual <- as.character(df$Bsmt.Qual)
df$Bsmt.Qual[df$Bsmt.Qual == "Ex"] <- "Excellent"
df$Bsmt.Qual[df$Bsmt.Qual == "Gd"] <- "Good"
df$Bsmt.Qual[df$Bsmt.Qual == "TA"] <- "Typical"
df$Bsmt.Qual[df$Bsmt.Qual == "Fa"] <- "Fair"
df$Bsmt.Qual[df$Bsmt.Qual == "Po"] <- "Poor"
df$Bsmt.Qual <- as.factor(df$Bsmt.Qual)
```


- Reorder the factor levels of two categorical variables: Kitchen_Qual and Bsmt_Qual: "Excellent", "Good", "Fair", "Typical", "Poor"
```{r}
df$Kitchen.Qual <- factor(df$Kitchen.Qual, levels=c("Excellent", "Good", "Fair", "Typical", "Poor"))
df$Bsmt.Qual <- factor(df$Bsmt.Qual, levels=c("Excellent", "Good", "Fair", "Typical", "Poor"))
```

- Replace contents of 4 columns: "Garage.Cars", "Garage.Area", "Total.Bsmt.SF", "Full.Bath"
```{r}
df$Garage.Cars[df$Garage.Cars == 0] <- 0.01
df$Garage.Area[df$Garage.Area == 0] <- 0.01
df$Total.Bsmt.SF[df$Total.Bsmt.SF == 0] <- 0.01
df$Full.Bath[df$Full.Bath == 0] <- 0.01
```


## 2.3. Data transformation

## 2.3.1. Convert categorical variables to dummies

- Create a set of categorical variable
```{r}
cate = c("MS.SubClass", "MS.Zoning", "Street", "Lot.Shape", "Land.Contour", "Utilities", "Lot.Config", "Land.Slope", "Neighborhood", "Condition.1", "Condition.2", "Bldg.Type", "House.Style", "Roof.Style", "Roof.Matl", "Exterior.1st", "Exterior.2nd", "Mas.Vnr.Type", "Exter.Qual", "Exter.Cond", "Foundation", "Bsmt.Qual", "Bsmt.Cond", "Bsmt.Exposure", "BsmtFin.Type.1", "BsmtFin.Type.2", "Heating", "Heating.QC", "Central.Air", "Electrical", "Kitchen.Qual", "Functional", "Fireplace.Qu", "Garage.Type", "Garage.Finish", "Garage.Qual", "Garage.Cond", "Paved.Drive", "Sale.Type", "Sale.Condition")
```

- convert categorical variables to dummies
```{r}
dfcate = df[,cate]
```

- The same code was used for all categorical variables
```{r}
for(level in unique(dfcate$MS.SubClass)){dfcate[paste("MS.SubClass", level, sep = "_")] <- ifelse(dfcate$MS.SubClass == level, 1, 0)}

for(level in unique(dfcate$MS.Zoning)){dfcate[paste("MS.Zoning", level, sep = "_")] <- ifelse(dfcate$MS.Zoning == level, 1, 0)}

for(level in unique(dfcate$Street)){dfcate[paste("Street", level, sep = "_")] <- ifelse(dfcate$Street == level, 1, 0)}

for(level in unique(dfcate$Lot.Shape)){dfcate[paste("Lot.Shape", level, sep = "_")] <- ifelse(dfcate$Lot.Shape == level, 1, 0)}

for(level in unique(dfcate$Land.Contour)){dfcate[paste("Land.Contour", level, sep = "_")] <- ifelse(dfcate$Land.Contour == level, 1, 0)}

for(level in unique(dfcate$Utilities)){dfcate[paste("Utilities", level, sep = "_")] <- ifelse(dfcate$Utilities == level, 1, 0)}

for(level in unique(dfcate$Lot.Config)){dfcate[paste("Lot.Config", level, sep = "_")] <- ifelse(dfcate$Lot.Config == level, 1, 0)}

for(level in unique(dfcate$Land.Slope)){dfcate[paste("Land.Slope", level, sep = "_")] <- ifelse(dfcate$Land.Slope == level, 1, 0)}

for(level in unique(dfcate$Neighborhood)){dfcate[paste("Neighborhood", level, sep = "_")] <- ifelse(dfcate$Neighborhood == level, 1, 0)}

for(level in unique(dfcate$Condition.1)){dfcate[paste("Condition.1", level, sep = "_")] <- ifelse(dfcate$Condition.1 == level, 1, 0)}

for(level in unique(dfcate$Condition.2)){dfcate[paste("Condition.2", level, sep = "_")] <- ifelse(dfcate$Condition.2 == level, 1, 0)}

for(level in unique(dfcate$Bldg.Type)){dfcate[paste("Bldg.Type", level, sep = "_")] <- ifelse(dfcate$Bldg.Type == level, 1, 0)}

for(level in unique(dfcate$House.Style)){dfcate[paste("House.Style", level, sep = "_")] <- ifelse(dfcate$House.Style == level, 1, 0)}

for(level in unique(dfcate$Roof.Style)){dfcate[paste("Roof.Style", level, sep = "_")] <- ifelse(dfcate$Roof.Style == level, 1, 0)}

for(level in unique(dfcate$MRoof.Matl)){dfcate[paste("Roof.Matl", level, sep = "_")] <- ifelse(dfcate$Roof.Matl == level, 1, 0)}

for(level in unique(dfcate$Exterior.1st)){dfcate[paste("Exterior.1st", level, sep = "_")] <- ifelse(dfcate$Exterior.1st == level, 1, 0)}

for(level in unique(dfcate$Exterior.2nd)){dfcate[paste("Exterior.2nd", level, sep = "_")] <- ifelse(dfcate$Exterior.2nd == level, 1, 0)}

for(level in unique(dfcate$Mas.Vnr.Type)){dfcate[paste("Mas.Vnr.Type", level, sep = "_")] <- ifelse(dfcate$Mas.Vnr.Type == level, 1, 0)}

for(level in unique(dfcate$Exter.Qual)){dfcate[paste("Exter.Qual", level, sep = "_")] <- ifelse(dfcate$Exter.Qual == level, 1, 0)}

for(level in unique(dfcate$Exter.Cond)){dfcate[paste("Exter.Cond", level, sep = "_")] <- ifelse(dfcate$Exter.Cond == level, 1, 0)}

for(level in unique(dfcate$Foundation)){dfcate[paste("Foundation", level, sep = "_")] <- ifelse(dfcate$Foundation == level, 1, 0)}

for(level in unique(dfcate$Bsmt.Qual)){dfcate[paste("Bsmt.Qual", level, sep = "_")] <- ifelse(dfcate$Bsmt.Qual == level, 1, 0)}

for(level in unique(dfcate$Bsmt.Cond)){dfcate[paste("Bsmt.Cond", level, sep = "_")] <- ifelse(dfcate$Bsmt.Cond == level, 1, 0)}

for(level in unique(dfcate$Bsmt.Exposure)){dfcate[paste("Bsmt.Exposure", level, sep = "_")] <- ifelse(dfcate$Bsmt.Exposure == level, 1, 0)}

for(level in unique(dfcate$BsmtFin.Type.1)){dfcate[paste("BsmtFin.Type.1", level, sep = "_")] <- ifelse(dfcate$BsmtFin.Type.1 == level, 1, 0)}

for(level in unique(dfcate$BsmtFin.Type.2)){dfcate[paste("BsmtFin.Type.2", level, sep = "_")] <- ifelse(dfcate$BsmtFin.Type.2 == level, 1, 0)}

for(level in unique(dfcate$Heating)){dfcate[paste("Heating", level, sep = "_")] <- ifelse(dfcate$Heating == level, 1, 0)}

for(level in unique(dfcate$Heating.QC)){dfcate[paste("Heating.QC", level, sep = "_")] <- ifelse(dfcate$Heating.QC == level, 1, 0)}

for(level in unique(dfcate$Central.Air)){dfcate[paste("Central.Air", level, sep = "_")] <- ifelse(dfcate$Central.Air == level, 1, 0)}

for(level in unique(dfcate$Electrical)){dfcate[paste("Electrical", level, sep = "_")] <- ifelse(dfcate$Electrical == level, 1, 0)}

for(level in unique(dfcate$Kitchen.Qual)){dfcate[paste("Kitchen.Qual", level, sep = "_")] <- ifelse(dfcate$Kitchen.Qual == level, 1, 0)}

for(level in unique(dfcate$Functional)){dfcate[paste("Functional", level, sep = "_")] <- ifelse(dfcate$Functional == level, 1, 0)}

for(level in unique(dfcate$Fireplace.Qu)){dfcate[paste("Fireplace.Qu", level, sep = "_")] <- ifelse(dfcate$Fireplace.Qu == level, 1, 0)}

for(level in unique(dfcate$Garage.Type)){dfcate[paste("Garage.Type", level, sep = "_")] <- ifelse(dfcate$Garage.Type == level, 1, 0)}

for(level in unique(dfcate$Garage.Finish)){dfcate[paste("Garage.Finish", level, sep = "_")] <- ifelse(dfcate$Garage.Finish == level, 1, 0)}

for(level in unique(dfcate$Garage.Qual)){dfcate[paste("Garage.Qual", level, sep = "_")] <- ifelse(dfcate$Garage.Qual == level, 1, 0)}

for(level in unique(dfcate$Garage.Cond)){dfcate[paste("Garage.Cond", level, sep = "_")] <- ifelse(dfcate$Garage.Cond == level, 1, 0)}

for(level in unique(dfcate$Paved.Drive)){dfcate[paste("Paved.Drive", level, sep = "_")] <- ifelse(dfcate$Paved.Drive == level, 1, 0)}

for(level in unique(dfcate$Sale.Type)){dfcate[paste("Sale.Type", level, sep = "_")] <- ifelse(dfcate$Sale.Type == level, 1, 0)}

for(level in unique(dfcate$Sale.Condition)){dfcate[paste("Sale.Condition", level, sep = "_")] <- ifelse(dfcate$Sale.Condition == level, 1, 0)}
```

## 2.3.2. Create a set of numerical variable
```{r}
numer = c("Lot.Frontage", "Lot.Area", "Overall.Qual", "Overall.Cond",
          "Year.Built", "Year.Remod.Add", "Mas.Vnr.Area", "BsmtFin.SF.1",
          "BsmtFin.SF.2", "Bsmt.Unf.SF", "Total.Bsmt.SF", "X1st.Flr.SF",
          "X2nd.Flr.SF", "Low.Qual.Fin.SF", "Gr.Liv.Area", "Bsmt.Full.Bath",
          "Bsmt.Half.Bath", "Full.Bath", "Half.Bath", "Bedroom.AbvGr",
          "Kitchen.AbvGr", "TotRms.AbvGrd", "Fireplaces", "Garage.Yr.Blt",
          "Garage.Cars", "Garage.Area", "Wood.Deck.SF", "Open.Porch.SF",
          "Enclosed.Porch", "X3Ssn.Porch", "Screen.Porch", "Pool.Area",
          "Misc.Val", "Mo.Sold", "Yr.Sold", "SalePrice")
dfnum = df[,numer]
```

## 2.3.3. Create a new dataframe combining numerical and dummy variables
```{r}
dfnew <- cbind(dfnum, dfcate[,41:297])
dim(dfnew)
```

## 2.4. Correlation analysis

- Build correlation atts with sale price
```{r}
correlations <- cor(dfnew)
price_cor <- correlations[,36]
summary(price_cor)
```

- correlation atts - price - Keep top 10 variables with the highest correlation
```{r}
head(price_cor[order(price_cor, decreasing = TRUE)],11)
```

- correlation mattrix
```{r}
top10 = c("Overall.Qual", "Gr.Liv.Area", "Garage.Cars", "Garage.Area", "X1st.Flr.SF", "Total.Bsmt.SF", "Full.Bath", "Year.Built", "Bsmt.Qual_Excellent", "Kitchen.Qual_Excellent","SalePrice")
dftop = dfnew[,top10]
dftop <- cor(dftop)
dftop
corrplot(dftop, na.label = "NA")
```

So we have 10 variables which belongs to belows types:
- Numeric: Overall.Qual, Gr.Liv.Area, Garage.Cars, Garage.Area, X1st.Flr.SF, Total.Bsmt.SF, Full.Bath , Year.Built
- Dummy varivales: Bsmt.Qual_Ex, Kitchen.Qual_Ex
- Categories to be considered:  Bsmt.Qual, Kitchen.Qual


# 3. Exploratory Data Analysis - EDA

## 3.1. Overall SalePrice
```{r}
g1 <- ggplot(df, aes(SalePrice))+ 
  geom_histogram(bins=60) + 
  labs(title = "Histogram of Saleprice") + 
  theme(plot.title=element_text(hjust=0.5)) +
  geom_vline(aes(xintercept=mean(SalePrice)), color="blue", linetype="dashed", size=1) +
  geom_text(aes(x=190000, label="Mean Value", y=250), colour="red", angle=90, vjust = 1.2, text=element_text(size=11))
g1
```

```{r}
ggplot(df, aes(sample=SalePrice)) +
  stat_qq(color="dodgerblue4") + 
  stat_qq_line(color="red") +
  scale_y_continuous(labels=function(y){y/10^6}) +
  labs(title="QQ Plot for SalePrices", y="Ordered Values") +
  theme(plot.title=element_text(hjust=0.5))
```

- Log transformation - SalePrice
```{r}
ggplot(df, aes(sample=log(SalePrice))) +
  stat_qq(color="dodgerblue4") + 
  stat_qq_line(color="red") +
  labs(title="QQ Plot for log(SalePrices)", y="Ordered values") +
  theme(plot.title=element_text(hjust=0.5))
```
- New Overall SalePrice
```{r}
ggplot(df, aes(log(SalePrice))) +
  geom_histogram(aes(y=..density..), 
                 bins=50, 
                 fill="dodgerblue4", 
                 alpha=0.6, 
                 color="white") +
  geom_density(adjust=3)
```

- Bsmt.Qual and Kitchen.Qual
```{r}
d1 <- ggplot(data.frame(df), aes(x=Bsmt.Qual)) +
  geom_bar()
d1

d2 <- ggplot(data.frame(df), aes(x=Kitchen.Qual)) +
  geom_bar()
d2
```

## 3.2. Plotting the marginal distributions of key categorical quantities of interest
- Bsmt.Qual and Kitchen.Qual
```{r}
d1 <- ggplot(data.frame(df), aes(x=Bsmt.Qual)) +
  geom_bar()
d1

d2 <- ggplot(data.frame(df), aes(x=Kitchen.Qual)) +
  geom_bar()
d2
```

- Bsmt.Qual and Kitchen.Qual in relationship with SalePrice
```{r}
b1 <- ggplot(df, aes(x = Bsmt.Qual, y = SalePrice)) +
    geom_boxplot(aes(color = factor(Bsmt.Qual)))
b1

b2 <- ggplot(df, aes(x = Kitchen.Qual, y = SalePrice)) +
    geom_boxplot(aes(color = factor(Kitchen.Qual)))
b2
```


## 3.3. Plotting the marginal distributions of key numerical quantities of interest
Overall.Qual, Gr.Liv.Area, Garage.Cars, Garage.Area, X1st.Flr.SF, Total.Bsmt.SF, Full.Bath, Year.Built
```{r}
p1 <- ggplot(df, aes(x=Overall.Qual)) + 
  geom_density()
p1 + geom_vline(aes(xintercept=mean(Overall.Qual)),
            color="blue", linetype="dashed", size=1)

p2 <- ggplot(df, aes(x=Gr.Liv.Area)) + 
  geom_density()
p2 + geom_vline(aes(xintercept=mean(Gr.Liv.Area)),
            color="blue", linetype="dashed", size=1)

p3 <- ggplot(df, aes(x=Garage.Cars)) + 
  geom_density()
p3 + geom_vline(aes(xintercept=mean(Garage.Cars)),
            color="blue", linetype="dashed", size=1)

p4 <- ggplot(df, aes(x=Garage.Area)) + 
  geom_density()
p4 + geom_vline(aes(xintercept=mean(Garage.Area)),
            color="blue", linetype="dashed", size=1)

p5 <- ggplot(df, aes(x=X1st.Flr.SF)) + 
  geom_density()
p5 + geom_vline(aes(xintercept=mean(X1st.Flr.SF)),
            color="blue", linetype="dashed", size=1)

p6 <- ggplot(df, aes(x=Total.Bsmt.SF)) + 
  geom_density()
p6 + geom_vline(aes(xintercept=mean(Total.Bsmt.SF)),
            color="blue", linetype="dashed", size=1)

p7 <- ggplot(df, aes(x=Full.Bath)) + 
  geom_density()
p7 + geom_vline(aes(xintercept=mean(Full.Bath)),
            color="blue", linetype="dashed", size=1)

p8 <- ggplot(df, aes(x=Year.Built)) + 
  geom_density()
p8 + geom_vline(aes(xintercept=mean(Year.Built)),
            color="blue", linetype="dashed", size=1)
```


## 3.4. Inspecting price against numerical variables of interest
Overall.Qual, Gr.Liv.Area, Garage.Cars, Garage.Area, X1st.Flr.SF, Total.Bsmt.SF, Full.Bath, Year.Built
```{r}
g1 <- ggplot(df, aes(factor(Overall.Qual), SalePrice, fill=factor(Overall.Qual))) + 
  geom_boxplot() +
  labs(x = "Overall.Qual", y = "SalePrice", title = "Boxplot of SalePrice vs. Overall.Qual") + 
  theme(legend.position = 'none', plot.title = element_text(size = 9, hjust=0.5))
g1

g2 <- ggplot(df, aes(factor(Gr.Liv.Area), SalePrice, fill=factor(Gr.Liv.Area))) + 
  geom_boxplot() +
  labs(x = "Gr.Liv.Area", y = "SalePrice", title = "Boxplot of SalePrice vs. Gr.Liv.Area") + 
  theme(legend.position = 'none', plot.title = element_text(size = 9, hjust=0.5))
g2

g3 <- ggplot(df, aes(factor(Garage.Cars), SalePrice, fill=factor(Garage.Cars))) + 
  geom_boxplot() +
  labs(x = "Garage.Cars", y = "SalePrice", title = "Boxplot of SalePrice vs. Garage.Cars") + 
  theme(legend.position = 'none', plot.title = element_text(size = 9, hjust=0.5))
g3

g4 <- ggplot(df, aes(factor( Garage.Area), SalePrice, fill=factor( Garage.Area))) + 
  geom_boxplot() +
  labs(x = " Garage.Area", y = "SalePrice", title = "Boxplot of SalePrice vs.  Garage.Area") + 
  theme(legend.position = 'none', plot.title = element_text(size = 9, hjust=0.5))
g4

g5 <- ggplot(df, aes(factor(X1st.Flr.SF), SalePrice, fill=factor(X1st.Flr.SF))) + 
  geom_boxplot() +
  labs(x = "X1st.Flr.SF", y = "SalePrice", title = "Boxplot of SalePrice vs. X1st.Flr.SF") + 
  theme(legend.position = 'none', plot.title = element_text(size = 9, hjust=0.5))
g5

g6 <- ggplot(df, aes(factor(Total.Bsmt.SF), SalePrice, fill=factor(Total.Bsmt.SF))) + 
  geom_boxplot() +
  labs(x = "Total.Bsmt.SF", y = "SalePrice", title = "Boxplot of SalePrice vs. Total.Bsmt.SF") + 
  theme(legend.position = 'none', plot.title = element_text(size = 9, hjust=0.5))
g6

g7 <- ggplot(df, aes(factor(Full.Bath), SalePrice, fill=factor(Full.Bath))) + 
  geom_boxplot() +
  labs(x = "Full.Bath", y = "SalePrice", title = "Boxplot of SalePrice vs. Full.Bath") + 
  theme(legend.position = 'none', plot.title = element_text(size = 9, hjust=0.5))
g7

g8 <- ggplot(df, aes(factor(Year.Built), SalePrice, fill=factor(Year.Built))) + 
  geom_boxplot() +
  labs(x = "Year.Built", y = "SalePrice", title = "Boxplot of SalePrice vs. Year.Built") + 
  theme(legend.position = 'none', plot.title = element_text(size = 9, hjust=0.5))
g8
```

## 3.5. Box-Cox transformation

```{r}
bc1 <- boxcox(lm(SalePrice~Overall.Qual, data=dfnew))
bc2 <- boxcox(lm(SalePrice~Gr.Liv.Area, data=dfnew))
bc3 <- boxcox(lm(SalePrice~Garage.Cars, data=dfnew))
bc4 <- boxcox(lm(SalePrice~Garage.Area, data=dfnew))
bc5 <- boxcox(lm(SalePrice~X1st.Flr.SF, data=dfnew))
bc6 <- boxcox(lm(SalePrice~Total.Bsmt.SF, data=dfnew))
bc7 <- boxcox(lm(SalePrice~Full.Bath, data=dfnew))
bc8 <- boxcox(lm(SalePrice~Year.Built, data=dfnew))
```

```{r}
optimal_lambda1 <- bc1$x[which(bc1$y == max(bc1$y))]
round(optimal_lambda1, 2)

optimal_lambda2 <- bc2$x[which(bc2$y == max(bc2$y))]
round(optimal_lambda2, 2)

optimal_lambda3 <- bc3$x[which(bc3$y == max(bc3$y))]
round(optimal_lambda3, 2)

optimal_lambda4 <- bc4$x[which(bc4$y == max(bc4$y))]
round(optimal_lambda4, 2)

optimal_lambda5 <- bc5$x[which(bc5$y == max(bc5$y))]
round(optimal_lambda5, 2)

optimal_lambda6 <- bc6$x[which(bc6$y == max(bc6$y))]
round(optimal_lambda6, 2)

optimal_lambda7 <- bc7$x[which(bc7$y == max(bc7$y))]
round(optimal_lambda7, 2)

optimal_lambda8 <- bc8$x[which(bc8$y == max(bc8$y))]
round(optimal_lambda8, 2)
```

All of 8 numerical variables are qualified to be transformed to log. But we need to take a look over their new relationships with SalePrice to decide which one should be used in log transformation.

## 3.6. New relationships with SalePrice after Box-Cox transformation
- log(Overall.Qual) with log(SalePrice)
```{r}
g1 <- ggplot(dfnew, aes(x=log(Overall.Qual), y=log(SalePrice))) +
  geom_point(color="dodgerblue4",size=0.7) + 
  labs(title="Log(SalePrice) vs. Log(Overall.Qual)") +
  geom_smooth(formula=y~x,method=lm, color="red") +
  theme(plot.title=element_text(hjust=0.5))

g2 <- ggplot(dfnew, aes(x=Overall.Qual, y=log(SalePrice))) +
  geom_point(color="dodgerblue4", size=0.7) + 
  labs(title="Log(SalePrice) vs. Overall.Qual") +
  geom_smooth(formula=y~x,method=lm, color="red") +
  theme(plot.title=element_text(hjust=0.5))

ggarrange(g1, g2, ncol=2,nrow=1)
```

- log(Gr.Liv.Area) with log(SalePrice)
```{r}
g1 <- ggplot(dfnew, aes(x=log(Gr.Liv.Area), y=log(SalePrice))) +
  geom_point(color="dodgerblue4",size=0.7) + 
  labs(title="Log(SalePrice) vs. Log(Gr.Liv.Area)") +
  geom_smooth(formula=y~x,method=lm, color="red") +
  theme(plot.title=element_text(hjust=0.5))

g2 <- ggplot(dfnew, aes(x=Gr.Liv.Area, y=log(SalePrice))) +
  geom_point(color="dodgerblue4", size=0.7) + 
  labs(title="Log(SalePrice) vs. Gr.Liv.Area") +
  geom_smooth(formula=y~x,method=lm, color="red") +
  theme(plot.title=element_text(hjust=0.5))

ggarrange(g1, g2, ncol=2,nrow=1)
```

- log(X1st.Flr.SF) with log(SalePrice)
```{r}
g1 <- ggplot(dfnew, aes(x=log(X1st.Flr.SF), y=log(SalePrice))) +
  geom_point(color="dodgerblue4",size=0.7) + 
  labs(title="Log(SalePrice) vs. Log(X1st.Flr.SF)") +
  geom_smooth(formula=y~x,method=lm, color="red") +
  theme(plot.title=element_text(hjust=0.5))

g2 <- ggplot(dfnew, aes(x=X1st.Flr.SF, y=log(SalePrice))) +
  geom_point(color="dodgerblue4", size=0.7) + 
  labs(title="Log(SalePrice) vs. X1st.Flr.SF") +
  geom_smooth(formula=y~x,method=lm, color="red") +
  theme(plot.title=element_text(hjust=0.5))

ggarrange(g1, g2, ncol=2,nrow=1)
```

- log(Year.Built) with log(SalePrice)
```{r}
g1 <- ggplot(dfnew, aes(x=log(Year.Built), y=log(SalePrice))) +
  geom_point(color="dodgerblue4",size=0.7) + 
  labs(title="Log(SalePrice) vs. Log(Year.Built)") +
  geom_smooth(formula=y~x,method=lm, color="red") +
  theme(plot.title=element_text(hjust=0.5))

g2 <- ggplot(dfnew, aes(x=Year.Built, y=log(SalePrice))) +
  geom_point(color="dodgerblue4", size=0.7) + 
  labs(title="Log(SalePrice) vs. Year.Built") +
  geom_smooth(formula=y~x,method=lm, color="red") +
  theme(plot.title=element_text(hjust=0.5))

ggarrange(g1, g2, ncol=2,nrow=1)
```

- log(Garage.Cars) with log(SalePrice)
```{r}
g1 <- ggplot(dfnew, aes(x=log(Garage.Cars), y=log(SalePrice))) +
  geom_point(color="dodgerblue4",size=0.7) + 
  labs(title="Log(SalePrice) vs. log(Garage.Cars)") +
  geom_smooth(formula=y~x,method=lm, color="red") +
  theme(plot.title=element_text(hjust=0.5))

g2 <- ggplot(dfnew, aes(x=Garage.Cars, y=log(SalePrice))) +
  geom_point(color="dodgerblue4", size=0.7) + 
  labs(title="Log(SalePrice) vs. Garage.Cars") +
  geom_smooth(formula=y~x,method=lm, color="red") +
  theme(plot.title=element_text(hjust=0.5))

ggarrange(g1, g2, ncol=2,nrow=1)
```

- log(Garage.Area) with log(SalePrice)
```{r}
g1 <- ggplot(dfnew, aes(x=log(Garage.Area), y=log(SalePrice))) +
  geom_point(color="dodgerblue4",size=0.7) + 
  labs(title="Log(SalePrice) vs. Log(Garage.Area)") +
  geom_smooth(formula=y~x,method=lm, color="red") +
  theme(plot.title=element_text(hjust=0.5))

g2 <- ggplot(dfnew, aes(x=Garage.Area, y=log(SalePrice))) +
  geom_point(color="dodgerblue4", size=0.7) + 
  labs(title="Log(SalePrice) vs. Garage.Area") +
  geom_smooth(formula=y~x,method=lm, color="red") +
  theme(plot.title=element_text(hjust=0.5))

ggarrange(g1, g2, ncol=2,nrow=1)
```

- log(Total.Bsmt.SF) with log(SalePrice)
```{r}
g1 <- ggplot(dfnew, aes(x=log(Total.Bsmt.SF), y=log(SalePrice))) +
  geom_point(color="dodgerblue4",size=0.7) + 
  labs(title="Log(SalePrice) vs. Log(Total.Bsmt.SF)") +
  geom_smooth(formula=y~x,method=lm, color="red") +
  theme(plot.title=element_text(hjust=0.5))

g2 <- ggplot(dfnew, aes(x=Total.Bsmt.SF, y=log(SalePrice))) +
  geom_point(color="dodgerblue4", size=0.7) + 
  labs(title="Log(SalePrice) vs. Total.Bsmt.SF") +
  geom_smooth(formula=y~x,method=lm, color="red") +
  theme(plot.title=element_text(hjust=0.5))

ggarrange(g1, g2, ncol=2,nrow=1)
```

- log(Full.Bath) with log(SalePrice)
```{r}
g1 <- ggplot(dfnew, aes(x=log(Full.Bath), y=log(SalePrice))) +
  geom_point(color="dodgerblue4",size=0.7) + 
  labs(title="Log(SalePrice) vs. Log(Full.Bath)") +
  geom_smooth(formula=y~x,method=lm, color="red") +
  theme(plot.title=element_text(hjust=0.5))

g2 <- ggplot(dfnew, aes(x=Full.Bath, y=log(SalePrice))) +
  geom_point(color="dodgerblue4", size=0.7) + 
  labs(title="Log(SalePrice) vs. Full.Bath") +
  geom_smooth(formula=y~x,method=lm, color="red") +
  theme(plot.title=element_text(hjust=0.5))

ggarrange(g1, g2, ncol=2,nrow=1)
```

As a result from this step, we decided to use log transformation for 4 independent numerical variables: Gr.Liv.Area, Garage.Area, X1st.Flr.SF, Total.Bsmt.SF. 

## 3.7. Colinearity

Regarding the Garage.Cars variable, we saw the colinearity between this variable and Garage.Area.
```{r}
# Investigate the relationship between Garage.Area and Garage.Cars
cor(dfnew$Garage.Area,dfnew$Garage.Cars)
cor(dfnew$Bsmt.Qual_Excellent,dfnew$Total.Bsmt.SF)

g1 <- ggplot(dfnew, aes(x=Garage.Area, y=Garage.Cars)) +
  geom_point(color="dodgerblue4",size=0.7) + 
  labs(title="Garage.Area vs. Garage.Cars") +
  geom_smooth(formula=y~x,method=lm, color="red") +
  theme(plot.title=element_text(hjust=0.5))
g1
```

-> Remove the Garage.Cars from the analysis

# 4. Modeling 
## 4.1. Modeling with top 9 variables

- Create training and teting dataset from dfnew 
```{r}
#Create Evaluation Sets
set.seed(1)
trainIndexes = sample(nrow(df),2000) 

#Creates training and test set from observations 
training = dfnew[trainIndexes,]
testing = dfnew[-trainIndexes,]
```

- Creat the first model including 7 numerical variables and 2 categorical variables:
```{r}
SalePrice_model1 <- lm(log(SalePrice) ~ Overall.Qual + log(Gr.Liv.Area) + log(Garage.Area) + log(X1st.Flr.SF) + log(Total.Bsmt.SF) +  Full.Bath + Year.Built + Kitchen.Qual_Excellent + Bsmt.Qual_Excellent, data = dfnew)
summary(SalePrice_model1)
```


- Create the second model by removing the Full.Bath varriable which had p-value > 0.05:
```{r}
SalePrice_model2 <- lm(log(SalePrice) ~ Overall.Qual + log(Gr.Liv.Area) + log(Garage.Area) + log(X1st.Flr.SF) + log(Total.Bsmt.SF) +  Year.Built + Kitchen.Qual_Excellent + Bsmt.Qual_Excellent, data = dfnew)
summary(SalePrice_model2)
```

- Try to investigate the interaction between log(Total.Bsmt.SF) and Bsmt.Qual:
```{r}
SalePrice_model3 <- lm(log(SalePrice) ~ Overall.Qual + log(Gr.Liv.Area) + log(Garage.Area) + log(X1st.Flr.SF) + log(Total.Bsmt.SF) +  Year.Built + Bsmt.Qual + log(Total.Bsmt.SF)*Bsmt.Qual, data = df)
summary(SalePrice_model3)
```

-> There is no insight into the model summary. Also, the Bsmt.Qual now became not significant.
-> The best model to fit: Model 2:
```{r}
summary(SalePrice_model2)
```

- Create a confidence interval for the model coefficients

```{r}
confint(SalePrice_model2, conf.level=0.95)
```

- Check the diagnostic plots for the model

```{r}
plot(SalePrice_model2)
```


## 4.2. Model with full variables 
- Create training and testing datasets
- Create Evaluation Sets
```{r}
dffull <- read.csv("dfnew.csv",
                     stringsAsFactors = FALSE, na.strings = c("NA", ""))

```

```{r}
set.seed(1)
trainIndexes = sample(nrow(dfnew),2000) 

#Creates training and test set from observations 
traindat = dffull[trainIndexes,]
testdat = dffull[-trainIndexes,]
```


```{r}
# Create a linear regression model
#Fit a linear regression model (to establish a familiar baseline)
model <- lm(SalePrice ~ . , traindat)

summary(model)

#Compute mean absolute error of model on train set
trainMAE <- mean(abs(model$residuals)) # 15081.7
trainMAE

#Calculate predictions for each observation in test set
predictY <- predict(model, testdat)

#Compute model residuals on the test set
testerrors <- testdat$SalePrice - predictY

#Compute mean absolute error of model on test set
testMAE <- mean(abs(testerrors)) #17257.71
testMAE

#Compute R^2
SSE <- sum((testdat$SalePrice- predictY)^2)
SST <- sum((testdat$SalePrice - mean(traindat$SalePrice))^2)
R_Square <- 1-(SSE/SST) 
R_Square# 0.9049922
```

```{r}
# back to new model 
traindat = dffull[trainIndexes,]
#The rest of the observations are set aside for test
testdat = dffull[-trainIndexes,]

model <- lm(SalePrice ~ . , traindat)

prevtrainMAE = trainMAE
for (i in 1:ncol(traindat)){
  # Create a linear regression model
  varimportance = data.frame(pvalue=coef(summary(model))[,4])
  varimportance$Predictor <- rownames(varimportance)
  #drop intercept
  varimportance = varimportance[2:nrow(varimportance),]
  #sort on pvalue
  varimportance = varimportance[order(-varimportance$pvalue),]
  print(nrow(varimportance))
  print(varimportance[1,1])
  if (varimportance[1,1]<.05){
    break
  }
  #drop highest p-value
  varimportance = varimportance[2:nrow(varimportance),]
  #store x variable names for use in lm()
  newxvalues = varimportance$Predictor
  #Fit a linear regression model
  model <- lm(SalePrice ~ . , traindat[,c("SalePrice",newxvalues)])
  
  #Compute mean absolute error of model on train set
  newtrainMAE <- mean(abs(model$residuals))
}
summary(model)

#lets look at the selected features
View(varimportance)

newtrainMAE  # 15555.18

#Calculate predictions for each observation in test set
predictY <- predict(model, testdat[,newxvalues])

#Compute model residuals on the test set
testerrors <- testdat$SalePrice - predictY

#Compute mean absolute error of model on test set
testMAE <- mean(abs(testerrors)) # 16545.16 
testMAE

#Compute R^2
SSE <- sum((testdat$SalePrice- predictY)^2)
SST <- sum((testdat$SalePrice - mean(traindat$SalePrice))^2)
R_Square <- 1-(SSE/SST) #0.9083282
R_Square
```


```{r}
###cross validation 
#Set up several data sets and Y variable vectors
#full
fullPriceY = dffull$SalePrice
dffull$SalePrice = NULL

#train
trainPriceY = traindat$SalePrice
traindat$SalePrice = NULL

#test
testPriceY = testdat$SalePrice
testdat$SalePrice = NULL

library(caret)
set.seed(1)
indx <- createFolds(trainPriceY, returnTrain = TRUE)
ctrl <- trainControl(method = "cv", index = indx)

set.seed(1)
lmTune0 <- train(x = traindat[,newxvalues], y = trainPriceY,
                 method = "lm",
                 trControl = ctrl)

lmTune0               

#Compute mean absolute error of model on train set
lmTune0$results$MAE # 16804.3

#Calculate predictions for each observation in test set
predictY <- predict(lmTune0, testdat[,newxvalues])

#Compute model residuals on the test set
testerrors <- testPriceY - predictY

#Compute mean absolute error of model on test set
testMAE <- mean(abs(testerrors)) # 16545.16
testMAE

#Compute R^2 

SSE <- sum((testPriceY- predictY)^2)
SST <- sum((testPriceY - mean(trainPriceY))^2)
R_Square <- 1-(SSE/SST) #0.9083282
R_Square
```
